---
title: "Collate Data"
output: html_notebook
---

# Goal 

Given a collection of datasets from GEO2R (stored as .txt files) convert them into a single data matrix of logFC values that can then be analyzed further (i.e. via clustering).

# Setup

## Libraries
```{r}
library(data.table)
library(psych)
```

## Inputs

Get a list of input files and read each as a data.frame.
```{r}
files <- list.files(pattern=".txt")
df_list <- lapply(files, fread)
```

# Wrangling

Collapse duplicated genes by taking median
```{r}
fc_list <- lapply(df_list, function(x) {
    cols <- c("Gene.symbol", "logFC")
    x[, cols, with=F]
})
fc_med <- lapply(fc_list, function(x) {
    x[, .(logFC=median(logFC)), .(Gene.symbol)]
})
```


Collect the files into a single data.frame, each time using the intersection of common genes.
```{r}
DF <- Reduce(function(x, y) {
    merge(x, y,
          by="Gene.symbol", no.dups=T)
}, fc_med)
# rename the columns
setattr(DF, "names",
        c("gene_symbol", paste("dataset", seq_along(files), sep="_")))
```

```{r}
# Save
fwrite(DF, "logFC_merged.txt", sep="\t")
```

```{r}
# cast as a matrix for plotting
M <- as.matrix(DF[, -1, with=F])
rownames(M) <- DF[[1]]
```

```{r}
pairs.panels(M, rug=F, pch=".", method="spearman")
```


