---
title: "Generate Networks"
output: html_notebook
---

# Goal

Given the background PPI and gene co-expression networks, convert them to graph objects that we can use as background networks. We'll use the `igraph` package to convert these interactions into a network object that we can compute on.

# Setup

## Libraries
```{r}
library(data.table)
library(igraph)
# our own custom functions
source("functions.R")
```

## Data
```{r}
# read the background network files
# these are simply 2-column tables with each row representing an interaction
# files have no header line
files <- list.files("background_networks", pattern=".csv", full.names=T)
network_dfs <- sapply(files, function(x) {
    fread(x, header=F)
}, USE.NAMES = T, simplify = F)
# update the names
names(network_dfs) <- gsub("background_networks/|\\.csv", "", names(network_dfs))
```


```{r}
# inputs from semantic similarity analysis - we can reuse the gene lists here
inputs <- readRDS("../semantic_similarity/inputs.rds")
```

# Convert to graph objects
```{r}
network_graphs <- sapply(network_dfs, function(x) {
   graph_from_data_frame(x, directed=F)
    }, USE.NAMES = T, simplify = F)
# count the number of edges - should be equal to number of rows in input data.frames
cbind(
    unlist(lapply(network_dfs, nrow)),
    unlist(lapply(network_graphs, function(x) length(E(x))))
)
```
Ok that looks good. Let's save it.

```{r}
# save
saveRDS(network_graphs, "background_networks.rds")
```

# Filter gene lists
```{r}
# filter each of the disease and drug gene lists so that it contains
# only genes in the graph object
# need to do this for each tissue separately
inputs_filt <- inputs[c("drug_targets", "disease_genes")]
universe <- names(V(g))
inputs_filt$disease_genes <- filter_gene_lists(inputs_filt$disease_genes,
                                               universe = universe)
inputs_filt$drug_targets <- sapply(inputs_filt$drug_targets, function(x) {
    filter_gene_lists(x, universe = universe)
}, USE.NAMES = T)
```




# Reformat into nested list

This will keep all our inputs in a single object and will simplify things for later analysis. We'll ignore the 'commonly expressed' genes in this analysis because we don't have a background network explicitly for them.

```{r}
# little helper function to organize the data
hf <- function(tissues, networks=network_graphs,
               disease_genes=inputs$disease_genes,
               drug_targets=inputs$drug_targets,
               patt="^PPI") {
    # initial list to store the data
    l <- sapply(tissues, function(x) {
        idx <- grep(paste(patt, x, sep="_"), names(networks))
        bn <- networks[[idx]]
        dg <- disease_genes[[x]]
        dt <- drug_targets[[x]]
        # filter so that the gene lists are all subsets of the genes in the
        # background network
        universe <- names(V(bn))
        dg <- intersect(dg, universe)
        dt <- filter_gene_lists(dt, universe = universe)
        list(
            background_network = bn,
            disease_genes = dg,
            drug_targets = dt
        )
    }, USE.NAMES = T, simplify = F)
    return(l)
}
```

```{r}
tissues <- setdiff(names(inputs$disease_genes), "Commonly_expressed")
obj <- list(
    PPI = hf(tissues, patt="^PPI"),
    GCN = hf(tissues, patt="^GCN")
)
```

```{r}
# save
saveRDS(obj, "inputs.rds")
```

