---
title: "Generate Networks"
output: html_notebook
---

# Goal

Use pre-computed background PPI and gene co-expression networks that have already been converted them to graph objects for analysis using drugbank gene lists. We'll reuse those same networks here with the LINCS gene sets. So, the only thing to do here is to filter the LINCS gene lists and generate an inputs object we can use for downstream analysis.


# Setup

## Libraries
```{r}
library(data.table)
library(igraph)
library(cmapR)
# our own custom functions
source("functions.R")
```

## Data

```{r}
# lincs gene lists as gmt files
up_sets <- parse_gmt("../LINCS_gene_sets/gene_sets_up.gmt")
dn_sets <- parse_gmt("../LINCS_gene_sets/gene_sets_dn.gmt")
```

```{r}
# ignore the up/down direction and concatenate into a single list for each
# drug
up_drugs <- Reduce(c, lapply(up_sets, function(x) x$desc))
dn_drugs <- Reduce(c, lapply(dn_sets, function(x) x$desc))
identical(up_drugs, dn_drugs) # this is true, they are the same lists
gene_lists <- sapply(up_drugs, function(x) {
    idx <- which(up_drugs == x)
    c(up_sets[[idx]]$entry, dn_sets[[idx]]$entry)
}, USE.NAMES = T, simplify = F)
```


```{r}
# pre-generated background networks
background_networks <- readRDS("background_networks.rds")
```

```{r}
# and inputs from prior analysis
inputs <- readRDS("inputs.rds")
```


# Filter gene lists

and save into an object we can use as input for downstream analysis
```{r}
# filter each of the LINCS gene lists so that it contains
# only genes in the graph object
# need to do this for each tissue separately
tissues <- unique(gsub("^PPI_|^GCN_", "", names(background_networks)))
network_types <- c("PPI", "GCN")
input_obj <- sapply(network_types, function(x) {
    networks <- background_networks[grep(x, names(background_networks))]
    gene_lists <- sapply(tissues, function(y) {
        bn <- networks[[grep(y, names(networks))]]
        universe <- names(V(bn))
        filt <- filter_gene_lists(gene_lists, universe = universe)
        list(
            background_network = bn,
            disease_genes = inputs[[x]][[y]]$disease_genes,
            gene_lists = filt
        )
    }, USE.NAMES = T, simplify = F)
}, USE.NAMES = T, simplify = F)
```


```{r}
# save
saveRDS(input_obj, "inputs_LINCS_backward_compat.rds", version = 2)
```

