---
title: "Proximity Score"
output: html_notebook
---

# Goal

Compute proximities between disease and drug gene lists, using the the interactome graph object. 

# Setup

## Libraries
```{r}
# library(data.table)
library(igraph)
library(parallel)
# our own custom functions
source("functions.R")
```

## Data
```{r}
# inputs from Generate_Networks.Rmd
inputs <- readRDS("inputs.rds")
```

# Analysis

## Generate proximity matrices

Within each tissue, for each diease gene list and each drug, compute the proximity score for each pair of genes. This will result in a matrix of disease genes x drug targets where each value is a proximity score. Note this will take some time, even with parallelization.
```{r}
prox_matrix_list <- sapply(inputs, function(assoc_type) {
    sapply(assoc_type, function(tissue) {
        bn <- tissue$background_network
        dg <- tissue$disease_genes
        dt <- tissue$drug_targets
        tmp <- mclapply(dt, function(dt_genes) {
            prox_mat <- proximity_matrix(bn, dg, dt_genes)
        }, mc.cores=4)
        names(tmp) <- names(dt)
        return(tmp)
    }, USE.NAMES = T, simplify = F)
}, USE.NAMES = T, simplify = F)
```

```{r}
# save this
saveRDS(prox_matrix_list, "prox_matrix_list.rds")
```

## Aggregate

Compute aggregate measures of proximity within each matrix, as described in Guney et al 2015 Nat. Comm. and wrap into a data.table.
```{r}
# apply summary function
prox_matrix_summ <- sapply(prox_matrix_list, function(assoc_type) {
    sapply(assoc_type, function(tissue) {
        sapply(tissue, function(drug) {
            summarize_prox_mat(drug)
        }, USE.NAMES = T, simplify = F)
    }, USE.NAMES = T, simplify = F)
}, USE.NAMES = T, simplify = F)
```

```{r}
# wrap up
ppi_df <- rbindlist(lapply(prox_matrix_summ$PPI, rbindlist, idcol="drug"), idcol="tissue")
gcn_df <- rbindlist(lapply(prox_matrix_summ$GCN, rbindlist, idcol="drug"), idcol="tissue")
prox_df <- rbindlist(list(PPI=ppi_df, GCN=gcn_df), idcol="assoc_type")
```

```{r}
# save
fwrite(prox_df, "proximity_scores.txt", sep="\t")
```

