---
title: "Proximity Null"
output: html_notebook
---

# Goal

Generate null distributions of proximity scores against which we can compare the true proximity scores to generate z-scores and/or p-values. The null will be comprised of size-matched groups of genes with similar degree distribution as the genes in the true comparison. 

# Setup

## Libraries
```{r}
# library(data.table)
library(igraph)
library(parallel)
# our own custom functions
source("functions.R")
```

## Data
```{r}
# inputs generated in Generate_Networks.Rmd
inputs <- readRDS("inputs.rds")
```

# Analysis

## Compute degree

Compute the degree of each node within each background network and append those values to the inputs object.
```{r}
tissues <- names(inputs$PPI)
for (x in tissues) {
    ppi_deg <- degree(inputs$PPI[[x]]$background_network)
    gcn_deg <- degree(inputs$GCN[[x]]$background_network)
    # quick look at the distributions
    hist(ppi_deg, main=paste("PPI", x), breaks=30)
    hist(gcn_deg, main=paste("GCN", x), breaks=30)
    inputs$PPI[[x]]$node_degree <- ppi_deg
    inputs$GCN[[x]]$node_degree <- gcn_deg
}

```
Looks like most are pretty low degree. 

Let's see whether there is a difference in degree distribution between the genes in each disease and drug target list. If not, then we don't need to account for this when generating the null.
```{r}
(deg_test_disease <- rbindlist(sapply(tissues, function(x) {
    compare_degree_distributions(inputs$PPI[[x]]$node_degree,
                                 inputs$PPI[[x]]$disease_genes)
    }, USE.NAMES = T, simplify = F),
    idcol="tissue"))
```

The disease genes seem to have different degree distributions from the entire background network, especially by the Wilcox test (rs in the table above). I think this test may be the most appropriate given its non-parametric nature and that the underlying node degree distributions are not normally distributed. So this does suggest that to be safe, we should account for degree distribution when computing the null.

## Bin based on degree

Next we'll bin the nodes into buckets based on degree.
```{r}
# use one global set of breaks. I think this should be ok.
# node with largest degree is 601
# use small bin size for better precision
breaks <- seq(0, 605, by=5) 
for (x in tissues) {
    inputs$PPI[[x]]$node_degree_binned <- cut(inputs$PPI[[x]]$node_degree,
                                              breaks = breaks,
                                              include.lowest = T)
    inputs$GCN[[x]]$node_degree_binned <- cut(inputs$GCN[[x]]$node_degree,
                                              breaks = breaks,
                                              include.lowest = T)
}
```


## Generate random gene lists

Then to generate the null, we'll need to sample in such a way that we accurately represent the degree distributions in the true gene lists. Do this for each network type and tissue separately.
```{r}
# first generate the sampled gene lists for disease genes
# this can take ~30 mins
set.seed(42)
nperm <- 1e3
networks <- c("PPI", "GCN")
null_lists <- sapply(networks, function(x) {
    sapply(tissues, function(y) {
        deg <- inputs[[x]][[y]]$node_degree
        deg_binned <- inputs[[x]][[y]]$node_degree_binned
        disease_genes <- inputs[[x]][[y]]$disease_genes
        disease_gene_lists <- lapply(1:nperm, function(i) {
            sample_degree(deg, deg_binned, disease_genes)
        })
        drug_targets <- inputs[[x]][[y]]$drug_targets
        drug_target_lists <- sapply(drug_targets, function(drug) {
            lapply(1:nperm, function(i) {
                sample_degree(deg, deg_binned, drug)
            })
        }, USE.NAMES = T, simplify = F)
        return(list(disease_genes = disease_gene_lists,
                    drug_targets = drug_target_lists))
    }, USE.NAMES = T, simplify = F)
}, USE.NAMES = T, simplify = F)
# save this result
saveRDS(null_lists, "null_lists.rds")
```

## Compute null proximities

Now that we have the null gene lists, we can compute the null proximity scores. Note that this will take some time, even with parallelization.
```{r}
null_prox_matrix_list <- 
    lapply(1:nperm, function(i) {
        sapply(networks, function(network) {
            sapply(tissues, function(tissue) {
                bn <- inputs[[network]][[tissue]]$background_network
                dg <- null_lists[[network]][[tissue]]$disease_genes[[i]]
                # extract the ith null list for each drug
                dt <- sapply(null_lists[[network]][[tissue]]$drug_targets,
                             function(x) x[[i]])
                tmp <- mclapply(dt, function(dt_genes) {
                    prox_mat <- proximity_matrix(bn, dg, dt_genes)
                }, mc.cores=12)
                names(tmp) <- names(dt)
                return(tmp)
            }, USE.NAMES = T, simplify = F)
    }, USE.NAMES = T, simplify = F)
})
# save this
saveRDS(null_prox_matrix_list, "null_prox_matrix_list.rds")
```

--- NOTE THIS SECTION IN FLUX DUE TO TIME COMPLEXITY ---
```{r}
# refactor to run each network/tissue on its own and save the intermediate results
for (network in networks2) {
    for (tissue in tissues3) { # edited to skip completed tissues
        # running out of memory at intermediaate steps,
        # so save the results in chunks
        chunk_size <- 100
        chunks <- chunk_vector(seq_len(nperm), chunk_size)
        for (y in seq_along(chunks)) {
            res <- lapply(chunks[[y]], function(i) {
                message(sprintf("%s %s chunk %d iter %d", network, tissue, y, i))
                bn <- inputs[[network]][[tissue]]$background_network
                dg <- null_lists[[network]][[tissue]]$disease_genes[[i]]
                # extract the ith null list for each drug
                dt <- sapply(null_lists[[network]][[tissue]]$drug_targets,
                             function(x) x[[i]])
                tmp <- mclapply(dt, function(dt_genes) {
                    prox_mat <- proximity_matrix(bn, dg, dt_genes)
                }, mc.cores=20)
                names(tmp) <- names(dt)
                return(tmp)
            })
            ofile <- sprintf("null_matrices_%s_%s_chunk_%d.rds", network, tissue, y)
            saveRDS(res, ofile)
        }
    }
}
```

## Aggregate null proximites

Because the computations are taking so long, we're running each network type/tissue combination on its own and saving the results as an individual .rds file. So we'll need to load each of these and aggregate individually, then collate the results.

Some of the null matrices were run in a chunked fashion, so we'll want to collate those first. The goal will be to have a single .rds file for each tissue / network type. This is an intermediate step that just to organize the files.
```{r}
chunked_files <- list.files("null_matrix_chunks", pattern="chunk.*\\.rds", full.names=T)
f <- gsub("_chunk_.*\\.rds", "", basename(chunked_files))
# f <- gsub("null_matrices_", "", f)
grp <- split(chunked_files, f)
```

```{r}
for (x in names(grp)) {
    message(x)
    files <- grp[[x]]
    collated <- Reduce(c, lapply(files, readRDS))
    saveRDS(collated, sprintf("%s.rds", x))
}
```

```{r}
# PPI skeletal muscle files are large, need to handle separately
sk1 <- Reduce(c, lapply(grp$null_matrices_PPI_Skeletal_muscle[1:5], readRDS))
saveRDS(sk1, "null_matrices_PPI_Skeletal_muscle_1.rds")
rm(sk1)
gc()
sk2 <- Reduce(c, lapply(grp$null_matrices_PPI_Skeletal_muscle[6:10], readRDS))
saveRDS(sk2, "null_matrices_PPI_Skeletal_muscle_2.rds")
rm(sk2)
gc()
```


```{r}
# null_prox_matrix_list <- list(
#     PPI = list(),
#     GCN = list()
# )
# null_prox_matrix_list$PPI[["Adipose_subcutaneous"]] <-
#     readRDS("null_matrices_PPI_Adipose_subcutaneous.rds")
```

Summarize by computing some aggregate stats on each individual proximity matrix. Need to do this on each .rds file separately due to memory constraints.

```{r}
null_files <- list.files(pattern="null_matrices", full.names = T)
```

```{r}
# NB these completed without modification - add them to list so we can skip
done <- c(
    "./null_matrices_GCN_Adipose_subcutaneous.rds",
    "./null_matrices_GCN_Adipose_visceral.rds",
    "./null_matrices_GCN_Liver.rds",
    "./null_matrices_GCN_Skeletal_muscle.rds"
)
for (nf in setdiff(null_files, done)) {
    message(nf)
    res <- readRDS(nf)
    agg <- sapply(res, function(perm) {
            # each tissue has a list of 1,000 permutations,
            # where each permutation is a list of matrices, one per drug
            sapply(perm, function(drug) {
                summarize_prox_mat(drug)
            }, simplify = F)
        }, USE.NAMES = T, simplify = F)
    df <- rbindlist(lapply(agg, rbindlist, idcol="drug"))
    # save the text file
    fname <- gsub("null_matrices_", "aggregated_", basename(nf))
    fname <- gsub("\\.rds$", "\\.txt", fname)
    fwrite(df, fname, sep="\t")
    rm(res)
    gc()
    # track those that are done
    done <- c(done, nf)
}
```

Collate all summarized stats for a given network / tissue / drug combination.
Here we want a data.frame that has all the iterations for every network / tissue / drug combination. We can then compute whatever aggregate stats we like on that. We can get this by collating the aggregated_* text files generated in the previous step.

```{r}
agg_files <- list.files(pattern="^aggregated_.*\\.txt$", full.names=T)
null_prox_df <- rbindlist(lapply(agg_files, function(x) {
    fname <- gsub("^aggregated_", "", basename(x))
    fname <- gsub("_[0-9]\\.txt$", "", fname)
    fname <- gsub("\\.txt$", "", fname)
    parts <- unlist(strsplit(fname, "_"))
    network <- parts[1]
    tissue <- paste(parts[-1], collapse="_")
    message(paste(network, tissue))
    tmp <- fread(x)
    tmp$network <- network
    tmp$tissue <- tissue
    return(tmp)
}))
# wrap up 
fwrite(null_prox_df, "null_prox_aggregated.txt", sep="\t")
```

Let's take a look at the distribution of mean_prox
```{r}
hist(null_prox_df$mean_prox)
```
Seems somewhat normal shaped, although there is a long tail to the right, but not too bad. Should be appropriate for z-scoring.

Compute some aggregate stats on the null distributions of mean_prosx. We need the mean and standard deviation so we can then compute z-scores
```{r}
null_prox_stats <- null_prox_df[, .(mean_prox = mean(mean_prox),
                                    sd_prox = sd(mean_prox)),
                                .(network, tissue, drug)]
# save
fwrite(null_prox_stats, "null_prox_stats_for_zscore.txt", sep="\t")
```

