---
title: "R Notebook"
output: html_notebook
---

This short snippet will illustrate how to parse and organize pathway gene lists stored in long format so that they can be used for downstream analysis.

Read existing `inputs` object from semantic similarity analysis. We will simply be modifying the `inputs$disease_genes` object, which is simply a list of vectors, where each vector contains the genes for a particular pathway.
```{r}
inputs <- readRDS("../semantic_similarity/inputs.rds")
```

Read first sheet of excel table. Modify this path as necessary.
```{r}
if (!require(readxl)) install.packages("readxl")
library(readxl)
library(data.table)

# modify this path as needed
pathways <- read_xlsx("~/work/Subhajit/Co-Expression_network/pathways_format.xlsx")
pathways <- data.table(pathways)
# convert the gene ids to characters
pathways$entrez_id <- as.character(pathways$entrez_id)
```

Split the long table of pathways into a list of lists using the `split` function
```{r}
pathway_lists <- split(pathways$entrez_id, pathways$pathways)
```

Now we can simply overwrite the `disease_genes` object with our new set of pathway lists.
```{r}
inputs$disease_genes <- pathway_lists
```

The following code is modified slightly because now we have a set of pathway gene lists, rather than a single disease gene list for each tissue type.
```{r}
# little helper function to organize the data
hf <- function(tissues, networks=network_graphs,
               disease_genes=inputs$disease_genes,
               drug_targets=inputs$drug_targets,
               patt="^PPI") {
    # initial list to store the data
    l <- sapply(tissues, function(x) {
        idx <- grep(paste(patt, x, sep="_"), names(networks))
        bn <- networks[[idx]]
        dg <- disease_genes # not specific to any tissue
        dt <- drug_targets[[x]]
        # filter so that the gene lists are all subsets of the genes in the
        # background network
        universe <- names(V(bn))
        dg <- filter_gene_lists(dg, universe = universe)
        dt <- filter_gene_lists(dt, universe = universe)
        list(
            background_network = bn,
            disease_genes = dg,
            drug_targets = dt
        )
    }, USE.NAMES = T, simplify = F)
    return(l)
}
```

```{r}
tissues <- setdiff(names(inputs$drug_targets), "Commonly_expressed")
obj <- list(
    PPI = hf(tissues, patt="^PPI"),
    GCN = hf(tissues, patt="^GCN")
)
```

```{r}
# save
saveRDS(obj, "inputs_pathway_lists.rds")
```

