---
title: "Zscore"
output: html_notebook
---

# Goal

Compute z-scores from proximity scores and null distributions. Could possibly also generate a network visualization by thresholding z-scores.

# Setup

## Libraries
```{r}
library(data.table)
library(igraph)
library(ggplot2)
library(ggrepel)
# library(parallel)
# our own custom functions
source("functions.R")
```

## Inputs
```{r}
# actual proximity scores
prox_df <- fread("proximity_scores_LINCS.txt")

# aggregated null stats
null_prox_df <- fread("null_prox_LINCS_stats_for_zscore.txt")
```

```{r}
# join
prox_merged <- merge(prox_df, null_prox_df,
                     by.x=c("assoc_type", "tissue", "drug"),
                     by.y=c("network", "tissue", "drug"),
                     suffixes = c("_actual", "_null"))
dim(prox_merged)
```

# Analysis

## z-scores (relative proximity)
```{r}
# compute z-scores
prox_merged[, prox_zs := (mean_prox_actual - mean_prox_null) / sd_prox,
            .(assoc_type, tissue, drug)]
```

```{r}
# save
fwrite(prox_merged, "proximity_zscores_LINCS.txt", sep="\t")
```


```{r}
# look at distribution by network, tissue type
ggplot(prox_merged, aes(x=prox_zs)) +
    geom_histogram() +
    xlab("proximity z-score") +
    ylab("frequency") +
    facet_grid(assoc_type ~ tissue, scales="free_y") +
    geom_vline(xintercept = 0, linetype="dashed", color="red")
```
Seems to have a longer negative tail, suggesting that there are some drug-gene proximities that are closer than you'd observe based on random chance. This is good. The Liver and 

```{r}
# look at the strongest associations
prox_merged <- prox_merged[order(prox_zs, decreasing=F)]
# prox_merged$tmp <- seq_len(nrow(prox_merged))
prox_merged[, tmp := seq_len(.N), .(assoc_type, tissue)]
p <- ggplot(prox_merged,
       aes(x=tmp, y=prox_zs)) +
    geom_point() +
    geom_label_repel(aes(label=drug), data=prox_merged[tmp <= 5]) +
    xlab("drugs") +
    ylab("proximity z-score") +
    facet_grid(assoc_type ~ tissue, scales="free_y")
# save as a PDF so we have more control over size
pdf("prox_zs_top5_LINCS.pdf", height=8.5, width=11)
print(p)
dev.off()
```


## Network viz

*NB* update to make more flexible for an arbitrary compound, or show the best compound from each network.

Visualize the significant associations in the background network by thresholding the z-scores.
```{r}
# get inputs object, which contains background networks
# and gene lists
inputs <- readRDS("inputs_LINCS_backward_compat_drugbank_filtered.rds")
```


```{r}
# get all vertices from network
verts <- V(inputs$PPI$Adipose_visceral$background_network)
```

Let's look at the subnetwork for a given compound and disease genes
```{r}
cp <- prox_merged[tissue=="Adipose_visceral"][which.min(prox_zs)]$drug
drug_targets <- inputs$PPI$Adipose_visceral$gene_lists[[cp]]
disease_genes <- inputs$PPI$Adipose_visceral$disease_genes
cp_idx <- which(names(verts) %in% drug_targets)
dx_idx <- which(names(verts) %in% disease_genes)
# extract a subgraph with these vertices plus nearest neighbors for context
vidx <- c(verts[cp_idx, .nei(1)], verts[dx_idx, .nei(1)])
subg <- induced_subgraph(inputs$PPI$Adipose_visceral$background_network, vidx)
```

```{r}
# make the plot
vcol <- rep("gray", length(vidx))
vcol[names(vidx) %in% drug_targets] <- "red"
vcol[names(vidx) %in% disease_genes] <- "blue"
vlab <- names(vidx)
vlab[!vlab %in% union(drug_targets, disease_genes)] <- ""
plot(subg,
     vertex.label=vlab,
     vertex.label.cex=0.5,
     vertex.label.dist=1,
     vertex.label.color="black",
     vertex.color=vcol,
     vertex.size=8,
     main=cp,
     sub="Adipose_visceral")
legend("bottomleft", legend = c("drug target", "disease gene"),
       col=c("red", "blue"), pch=19)
```


