---
title: "Semantic Similarity"
output: html_notebook
---

# Overview

The objective of this analysis is to compute the semantic similarity between tissue-specific lists of drug targets and disease genes. We'll do this using the R package [GOSemSim](https://bioconductor.org/packages/devel/bioc/vignettes/GOSemSim/inst/doc/GOSemSim.html#gene-semantic-similarity-measurement). And the corresponding paper [here.](https://academic.oup.com/bioinformatics/article/26/7/976/213143). Wang's similarity method, which I've chosen to use, is explained [here.](https://academic.oup.com/bioinformatics/article/23/10/1274/197095#1953560)

# Setup

## Libraries
```{r}
library(AnnotationHub)
library(GOSemSim)
library(data.table)
library(cmapR)
library(parallel)
library(ggplot2)
library(ggrepel)
source("functions.R")
```

```{r}
# make sure org.Hs.eg.db annotation package is installed, as GOSemSim needs
# this to compute similarities
if (!require(org.Hs.eg.db)) {
    BiocManager::install("org.Hs.eg.db")
    library(org.Hs.eg.db)
}
```

## Ontology

Restrict ontology to molecular functions. From the paper above:

"The GO used in measurement can be restricted by assigning the corresponding parameter to 'BP' (biological process), 'MF' (molecular function) and 'CC' (cellular component)."

```{r}
# Initialize a GOSemSimDATA object to contain the GO term to symbol mapping
hsGO <- godata('org.Hs.eg.db', ont="MF", keytype = "SYMBOL")
```


## Inputs

```{r}
# read from previously saved inputs object created in 'SemSim_Timing.Rmd'
inputs <- readRDS("inputs.rds")
```

### Update to include LINCS gene lists instead of drug bank lists

```{r}
# lincs gene lists as gmt files
up_sets <- parse_gmt("~/work/Subhajit/LINCS_gene_sets/gene_sets_up.gmt")
dn_sets <- parse_gmt("~/work/Subhajit/LINCS_gene_sets/gene_sets_dn.gmt")
```

```{r}
# ignore the up/down direction and concatenate into a single list for each
# drug
up_drugs <- Reduce(c, lapply(up_sets, function(x) x$desc))
dn_drugs <- Reduce(c, lapply(dn_sets, function(x) x$desc))
identical(up_drugs, dn_drugs) # this is true, they are the same lists
gene_lists <- sapply(up_drugs, function(x) {
    idx <- which(up_drugs == x)
    c(up_sets[[idx]]$entry, dn_sets[[idx]]$entry)
}, USE.NAMES = T, simplify = F)
```

```{r}
inputs$drug_targets <- NULL
inputs$gene_lists <- gene_lists
```

```{r}
# save
saveRDS(inputs, "inputs_LINCS_backward_compat.rds", ver=2)
```

```{r}
# and make a filtered version containing only those drugs that have drugbank data
# can use the proximity inputs to help with this
prox_inputs <- readRDS("../proximity/inputs_LINCS_backward_compat_drugbank_filtered.rds")
drug_bank_drugs <- names(prox_inputs$PPI$Liver$gene_lists)
inputs_filt <- inputs
inputs_filt$gene_lists <- inputs$gene_lists[drug_bank_drugs]
```

```{r}
# save
saveRDS(inputs_filt, "inputs_LINCS_backward_compat_drugbank_filtered.rds", ver=2)
```

# Computation

## Sanity check

As a sanity check, let's compute the similarity between the target lists for all the drugs in a given tissue. We should observe that drugs with similar mechanism cluster together.
```{r}
simtest <- mclusterSim(inputs$drug_targets$Commonly_expressed,
                       semData=hsGO, measure="Jiang")
d <- as.dist(1 - simtest)
hc <- hclust(d)
plot(hc)
```
Ok, this looks good. The estradiols, glitazones, and heavy metal drugs all form their own clusters.

## Drug to disease gene lists

We'll compute the similarity between each drug list and each tissue gene list, and wrap up the results into a matrix. Note that this takes a while, so one may want to simply load the .rds file with the results if available. For Windows users, mc.cores = 1 can be used. 
```{r}
tissues <- names(inputs$disease_genes)
tissue_res <- lapply(tissues, function(x) {
    disease_list <- inputs$disease_genes[[x]]
    drug_lists <- inputs$gene_lists
    disease_drug_sim(disease_list, drug_lists, semData=inputs$hsGO, measure="Jiang")
})
names(tissue_res) <- tissues
```

```{r}
# save since this took a while
saveRDS(tissue_res, "tissue_res_Jiang_LINCS.rds")
```


----- This is after the null has been run ------

Read in p-values and collate into a table for each tissue.
```{r}
pval_list <- readRDS("pval_lists_all_tissues_nperm_100_Jiang.rds")
```

```{r}
tissue_tables <- sapply(tissues, function(x) {
    res <- tissue_res[[x]]
    pvals <- pval_list[[x]]
    idx1 <- match(names(res), names(pvals))
    idx2 <- match(names(res), names(tissue_res_Jiang$drug_targets[[x]]))
    data.table(drug = names(res),
               # number of genes in the disease and drug lists
               ngene_disease = length(tissue_res_Jiang$disease_genes[[x]]),
               ngene_drug = lengths(tissue_res_Jiang$drug_targets[[x]][idx2]),
               semsim = res,
               pval = pvals[idx1])
}, USE.NAMES = T, simplify = F)
```

```{r}
# plot score vs. p-value for each tissue
for (x in tissues) {
    tmp <- tissue_tables[[x]]
    p <- ggplot(tmp, aes(x=semsim, y=pval)) +
        scale_y_reverse() +
        geom_point() +
        ggtitle(x)
        # geom_label_repel(aes(label=drug), data=tmp[pval <= 0.05])
    print(p)
}
```


Collate into a single large table. Include NA values for drugs that did not have a gene signature in a given tissue.
```{r}
# update column names so we can merge
names_to_change <- c("semsim", "pval", "ngene_disease", "ngene_drug")
for (x in tissues) {
    setnames(tissue_tables[[x]], names_to_change,
             paste(x, names_to_change, sep="_"))
}
```

```{r}
# merge them all together, keeping all entries from each table (all = T)
res_df <- Reduce(function(x, y) {
    merge(x, y, all=T)
}, tissue_tables)
```

```{r}
# save
fwrite(res_df, "results_table_collated_nperm_100_Jiang.txt", sep="\t")
```





